---
title: "Data processing & visualization script for scallop gsi paper"
author: "Lucy Williams & Anya Hopple"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load necessary libraries
library(tidyverse)
library(ggpubr)
library(patchwork) 
library(hrbrthemes) 
library(ggthemes)

# Set a custom theme for ggplot
theme_set(theme_tufte())
theme_update(plot.title =element_text(hjust=0.5),
             axis.line.x = element_line(color="black", size=1), 
             element_line(color="black", size=1),
             text = element_text(size=20),
             axis.text = element_text(size=15)) 
```

## Data import

```{r}

# Load the data set
datlarge <- read_csv("GSI.tidy.csv")
```

## Data cleaning and transformation

```{r}

# Data transformation
datlarge <- datlarge %>%
  mutate(
    # Setting categorical variables as factors
    Site = as.factor(Site),
    Gear_Type = as.factor(Gear_Type),
    Level_Position = as.factor(Level_Position),
    Sex = as.factor(Sex),
    Meat_Color = as.factor(Meat_Color),
    Year = as.factor(Year),
    # Setting Date as date
    Date = as.Date(Date),
    # Creating a new variable "type" based on the condition "Gear_Type";
    # if Lantern_Net or Cage, type = Farm | if not, type = Wild
    type = if_else(Gear_Type %in% c("Lantern_Net", "Cage"), "Farm", "Wild"),
    type = as.factor(type)
  )

# Data cleaning
# datlarge includes data from 2019-2022 & all gear types
# dat removes (1) 2019 data, (2) cage data, & (3) 1 week of mis-entered data in 2022

# Filtering out above data types
datlarge %>%
  filter(Year != 2019, 
         Gear_Type != "Cage",
         Date != "2022-06-28") ->
  dat

# Define shell size classes
dat$sc <- cut(dat$Shell_Height,
              breaks = c(30, 50, 60, 70, 80, 90, 100, 110, Inf),
              labels = c("<50", "50-59", "60-69", "70-79", "80-89", "90-99", "100-109", ">110"),
              include.lowest = FALSE)

dat$sc <- as.factor(dat$sc)
summary(dat$sc)
```

## Normality & statistical tests

```{r}
# Define the variables to test
variables_to_test <- c("GSI", "Meat_Weight", "Total_Viscera_Weight", "Gonad_Weight",
                       "Shell_Weight", "Shell_Height")

# Create lists to store plots and test results
density_qq_plots <- list()
shapiro_test_results <- list()

# Loop through each variable and create plots and perform Shapiro-Wilk test
for (var in variables_to_test) {
  
  # Density plot
  density_plot <- ggplot(dat, aes(x = .data[[var]])) +
    geom_density(stat = "density", alpha = 0.2) +
    xlab(paste("mean(", var, ")")) +
    ylab("Count")
  
  # QQ plot
  qq_plot <- ggqqplot(dat[[var]])
  
  # Add density and QQ plots to the list
  density_qq_plots[[var]] <- list(density_plot, qq_plot)
  
  # Shapiro-Wilk test
  shapiro_test_result <- shapiro.test(dat[[var]])
  shapiro_test_results[[var]] <- shapiro_test_result
}

# Create a separate table for Shapiro-Wilk test results
shapiro_test_table <- data.frame(
  Variable = variables_to_test,
  P_Value = sapply(shapiro_test_results, function(result) result$p.value)
)

# Suppress warning from ggarrange
suppressWarnings({
  # Print the plots using ggarrange
  library(ggpubr)
  ggarrange(
    density_qq_plots$GSI[[1]], density_qq_plots$GSI[[2]],
    density_qq_plots$Meat_Weight[[1]], density_qq_plots$Meat_Weight[[2]],
    density_qq_plots$Total_Viscera_Weight[[1]], density_qq_plots$Total_Viscera_Weight[[2]],
    density_qq_plots$Gonad_Weight[[1]], density_qq_plots$Gonad_Weight[[2]],
    density_qq_plots$Shell_Weight[[1]], density_qq_plots$Shell_Weight[[2]],
    density_qq_plots$Shell_Height[[1]], density_qq_plots$Shell_Height[[2]],
    nrow = 6, ncol = 2
  )
})

# Print the Shapiro-Wilk test results as a table
print(shapiro_test_table)
```

##Logged normality and statistical tests

```{r}
# Define the variables to test
variables_to_test <- c("GSI", "Meat_Weight", "Total_Viscera_Weight", "Gonad_Weight",
                       "Shell_Weight", "Shell_Height")

# Create lists to store plots and test results
density_qq_plots_log <- list()
shapiro_test_results_log <- list()

# Loop through each variable and create plots and perform Shapiro-Wilk test
for (var in variables_to_test) {
  
  # Density plot
  density_plot_log <- ggplot(dat, aes(x = log(.data[[var]]))) +
    geom_density(stat = "density", alpha = 0.2) +
    xlab(paste("mean(", var, ")")) +
    ylab("Count")
  
  # QQ plot
  qq_plot_log <- ggqqplot(log(dat[[var]]))
  
  # Add density and QQ plots to the list
  density_qq_plots_log[[var]] <- list(density_plot_log, qq_plot_log)
  
  # Shapiro-Wilk test
  shapiro_test_result_log <- shapiro.test(log(dat[[var]]))
  shapiro_test_results_log[[var]] <- shapiro_test_result_log
}

# Create a separate table for Shapiro-Wilk test results
shapiro_test_table_log <- data.frame(
  Variable = variables_to_test,
  P_Value = sapply(shapiro_test_results_log, function(result) result$p.value)
)

# Suppress warning from ggarrange
suppressWarnings({
  # Print the plots using ggarrange
  library(ggpubr)
  ggarrange(
    density_qq_plots_log$GSI[[1]], density_qq_plots_log$GSI[[2]],
    density_qq_plots_log$Meat_Weight[[1]], density_qq_plots_log$Meat_Weight[[2]],
    density_qq_plots_log$Total_Viscera_Weight[[1]], density_qq_plots_log$Total_Viscera_Weight[[2]],
    density_qq_plots_log$Gonad_Weight[[1]], density_qq_plots_log$Gonad_Weight[[2]],
    density_qq_plots_log$Shell_Weight[[1]], density_qq_plots_log$Shell_Weight[[2]],
    density_qq_plots_log$Shell_Height[[1]], density_qq_plots_log$Shell_Height[[2]],
    nrow = 6, ncol = 2
  )
})

# Print the Shapiro-Wilk test results as a table
print(shapiro_test_table)
```

## ANOVA

```{r}
# Define the variables to test
variables_to_test <- c("GSI", "Meat_Weight", "Total_Viscera_Weight", "Gonad_Weight", "Shell_Weight", "Shell_Height")

# Define the size class levels of interest
size_class_levels_of_interest <- c("80-89", "90-99", "100-109")


# Create a data frame to store the ANOVA results
anova_results_table <- data.frame(
  Size_Class = character(),
  Variable = character(),
  P_Value = numeric(),
  stringsAsFactors = FALSE
)

# Loop through each size class
for (size_class_level in size_class_levels_of_interest) {
  # Subset the data for the current size class
  subset_data <- dat[dat$sc == size_class_level, ]
  
  # Loop through each variable
  for (var in variables_to_test) {
    # Perform ANOVA analysis with the current [logged] variable as the dependent variable 
    anova_result <- summary(aov(log(subset_data[[var]]) ~ type, data = subset_data))
    
    # Extract the p-value from the ANOVA result
    p_value <- anova_result[[1]]$`Pr(>F)`[1]
    
    # Add the results to the data frame
    anova_results_table <- rbind(anova_results_table, data.frame(
      Size_Class = as.character(size_class_level),
      Variable = var,
      P_Value = p_value
    ))
  }
}

# Print the ANOVA results table
print(anova_results_table)
```

## ANOVA (Between size classes in Farm/Wild), Weeks 32:38
```{r}
# Define the variables to test
variables_to_test <- c("GSI", "Meat_Weight", "Total_Viscera_Weight", "Gonad_Weight", "Shell_Weight", "Shell_Height")

# Define the two treatments to run the loop through
type_levels_of_interest <- c("Farm", "Wild")

# Create a data frame to store the ANOVA results
anova_results_table <- data.frame(
  Type = character(),
  Variable = character(),
  P_Value = numeric(),
  stringsAsFactors = FALSE
)

# Loop through each size type
for (type_level in type_levels_of_interest) {
  # Subset the data for the current type
  subset_data <- dat %>% filter(Week %in% c(32:38), type == type_level)
  
  # Loop through each variable
  for (var in variables_to_test) {
    # Perform ANOVA analysis with the current [logged] variable as the dependent variable 
    anova_result <- summary(aov(log(subset_data[[var]]) ~ sc, data = subset_data))
    
    # Extract the p-value from the ANOVA result
    p_value <- anova_result[[1]]$`Pr(>F)`[1]
    
    # Add the results to the data frame
    anova_results_table <- rbind(anova_results_table, data.frame(
      Type = as.character(type_level),
      Variable = var,
      P_Value = p_value
    ))
  }
}

# Print the ANOVA results table
print(anova_results_table)

#I am unsure if this looks correct# 
```

## Variance

```{r}
#Online source for code/library : https://stat-methods.com/home/levenes-test-for-homogeneity-of-variance/

#library for test
library(car)

#Define variables to test
variables_to_test <- c("GSI", "Meat_Weight", "Total_Viscera_Weight", "Gonad_Weight", "Shell_Weight", "Shell_Height")

#Test variables in two treatment groups: wild and farmed

var_results_table <- data.frame(
  Variable = character(),
  F_value = numeric(),
  P_Value = numeric(),
  stringsAsFactors = FALSE
)

  
# Loop through each variable
for (var in variables_to_test) {
    # Perform Levene test with [logged] variables as the dependent variable 
    var_result <- leveneTest(log(dat[[var]]) ~ type, data = dat)
    
    p_value <- var_result$`Pr(>F)`
    f_value <- var_result$`F value`
    
    var_results_table <- rbind(var_results_table, data.frame(
      Variable = var,
      F_value = f_value,
      P_Value = p_value
    ))
    
   
    
  }

print(var_results_table)
#I do not know why it is printing twice - for AH#

#Variances are not equal
```


## T-test
```{r}
dat_test<- dat %>% rename(
  "Adductor Mass" = "Meat_Weight" ,
  "Total Viscera Mass" = "Total_Viscera_Weight" ,
  "Gonad Mass" = "Gonad_Weight", 
  "Shell Mass" = "Shell_Weight" ,
  "Shell Height" = "Shell_Height"
)
  
  
# Define the variables to test
variables_to_test<- c("GSI", "Adductor Mass",
           "Total Viscera Mass", "Gonad Mass", 
           "Shell Mass", "Shell Height")
# Define the size class levels of interest
size_class_levels_of_interest <- c("80-89", "90-99", "100-109")

#Dataframe for t.test results
t_test_results_table <- data.frame(
  Size_class = character(),
  Variables = character(),
  P_value = numeric(),
  stringsAsFactors = FALSE
)


# Loop through each size class
for (size_class_level in size_class_levels_of_interest) {
  # Subset the data for the current size class
  subset_data <- dat_test %>% filter(Week %in% c(32:38), sc == size_class_level)
  
  # Loop through each variable
  for (var in variables_to_test){
    t_test_result <- t.test(log(subset_data[[var]]) ~ type, data = subset_data, var.equal = FALSE)
    
    p.value <- t_test_result$p.value
    
    t_test_results_table <- rbind(t_test_results_table, data.frame(
      Size_class = as.character(size_class_level),
      Variables = var,
      p.value = p.value
    ))
  }
  
}
#library#
library(rstatix)
#P.value star significance table for plots
significance_table <- t_test_results_table %>% add_significance("p.value")

# Print the t-test results table

print(t_test_results_table)

#Pring with significance 
print(significance_table)



```

##Figure 1: Boxplots

```{r}
#Box plot of each variable separated farm/wild and size class

#Size class levels defined earlier

#Change variable names for future plotting 
dat_plot <- dat %>% rename(
  "Adductor Mass" = "Meat_Weight" ,
  "Total Viscera Mass" = "Total_Viscera_Weight" ,
  "Gonad Mass" = "Gonad_Weight", 
  "Shell Mass" = "Shell_Weight" ,
  "Shell Height" = "Shell_Height"
)
#library for scaling
library(scales)

#List to store plots
boxplot_list <- list()

#Selecting variables to plot
variables_to_plot <- c("GSI", "Adductor Mass", 
           "Total Viscera Mass", "Gonad Mass", 
           "Shell Mass", "Shell Height")

  
  # Loop through each variable
  for (var in variables_to_plot) {
    
    # Subset the data for the current size class
    subset_data <- dat_plot %>%
      filter(Week %in% c(32:38) & sc %in% c("80-89", "90-99", "100-109"))
                                 
    box_plot <- ggplot(data = subset_data, aes(y = log(.data[[var]]))) + 
      geom_boxplot(aes(x = type, fill = type)) + 
      facet_wrap(~sc, ncol = 4) + 
      scale_fill_grey(name = "", 
                      start = 0.3,
                      end = 1.0) +
      xlab("") + ylab(paste0("ln(", var, ")"))+
      scale_y_continuous(labels = label_number(accuracy = 1))+
      theme(legend.position="none") 
    
    boxplot_list[[var]] <- list(box_plot)
    
    
}

 boxplot_list 

  
suppressWarnings({
  ggarrange(
    boxplot_list$`Total Viscera Mass`[[1]],
    boxplot_list$`Shell Mass`[[1]],
    boxplot_list$`Adductor Mass`[[1]],
    boxplot_list$`Gonad Mass`[[1]],
    boxplot_list$GSI[[1]])
})


```
## Figure 2: Parsons & Dadswell plots

```{r}
#Load paper data
dads <- read_csv("dads_sum.csv") 

#The following code was used to summarize our data and then combined with the Parsons paper

#dat %>% filter(Week %in% 32:38) %>%
  #group_by(type, sc) %>% 
  #summarize(mean_gsi = mean(GSI, na.rm = TRUE),
            #n = length(GSI),
            #SD = sd(GSI, na.rm = TRUE), 
            #SE = sd(GSI, na.rm=TRUE) / sqrt(length(GSI)))

#Factoring for order of bar plots
dads$type <- factor(dads$type, levels = c("Farm", "Wild", "Parson et al"))
#Single plot

dads %>%
  ggplot(aes(x = sc, y = mean_gsi, fill = type)) + 
  geom_bar(stat = "identity", position = "dodge", color= "black") +
  xlab("Size Category (mm)")+
  ggtitle("") + ylab("GSI (%)") +
  geom_errorbar(aes(ymin = mean_gsi - SE, ymax = mean_gsi + SE), 
                width = 0.2, position = position_dodge(0.9)) +
  scale_x_discrete(limits = c("60-69", "70-79", "80-89","90-99", "100-109")) +
  scale_fill_grey(name = "", start = 0.2, end = 0.9) +theme(legend.position="bottom") +
  geom_text(aes(label = N, y = 0.9), position = position_dodge(0.9), size = 3.5, color = "white")




```

## Figure 3: Condition Indexes
```{r}

#Create total mass column = total viscera mass + shell mass

dat_plot <- dat_plot %>% mutate("Total Mass" = `Total Viscera Mass` + `Shell Mass` )

#Select variables to plot

variables_to_plot <- c("GSI", "Adductor Mass", 
           "Total Viscera Mass", "Gonad Mass", 
           "Shell Mass")

#Define plot list
index_plot_list <- list()

#Loop through each variable, not including shell height
  for (var in variables_to_plot) {
    
    # Subset the data for the current size class
    subset_data <- dat_plot %>%
      filter(Week %in% c(32:38) & sc %in% c("80-89", "90-99", "100-109"))
    
    index_plot <- ggplot(data = subset_data, 
                         aes(x = ,log(`Total Mass`), y = log(.data[[var]]), color = type))+ geom_point(shape = 1) + 
      scale_color_manual(values = c("cyan4", "orange")) +
      geom_smooth(aes(group = type, lty = type), method = "lm", color = "black" ) +
      stat_regline_equation(aes(label = ..rr.label..)) + ylim(c(1,5))+  xlim(3,6) +  
      ggtitle("",paste0(var, "(g)")) + ylab(paste0("ln(", var, ")")) + xlab("ln(Total Mass)")
    
    index_plot_list[[var]] <- list(index_plot)

    
}



suppressWarnings({
  ggarrange(
    index_plot_list$`Total Viscera Mass`[[1]],
    index_plot_list$`Shell Mass`[[1]],
    index_plot_list$`Adductor Mass`[[1]],
    index_plot_list$`Gonad Mass`[[1]],
    index_plot_list$GSI[[1]])
  })



```

## Figure 4: Time series of morphometric data
```{r}
#Pivot data frame to plot multiple morphometric data
dat_piv <- dat_plot %>% 
  pivot_longer(
    c(13:16, 18), 
    names_to = "morphometric", 
    values_to = "value")
#Plot data 
dat_piv %>% filter(Week %in% c(20:44)) %>% 
  ggplot(aes(x = Week, y = `value`, color = morphometric)) + 
  geom_smooth() + 
  facet_wrap(~type) + xlim(c(20,50)) +
  ylab("Mass")+
  scale_x_continuous(
    name = "Month", 
    breaks = c(23,27,32,36, 40), 
    labels = c("Jun", "Jul", "Aug", "Sep", "Oct")) +
  scale_color_discrete(name = "Morphometric")
                       

#Odd that total viscera and shell mass plot so similarly for Farm
#Checking out indiviual years to see if the trend remains
dat_plot %>% filter(Week %in% c(20:44)) %>% filter(type == "Farm") %>%
  ggplot(aes(x = Week)) +
  geom_point(aes(y =`Shell Mass`), color = "blue", alpha = 0.1) +
  geom_point(aes(y = `Total Viscera Mass`), color = "red", alpha = 0.1) +
  geom_smooth(aes(y =`Shell Mass`), color = "blue") + 
  geom_smooth(aes(y = `Total Viscera Mass`), color = "red") +
  facet_wrap(~Year)+
  ylab("Mass")+
  scale_x_continuous(
    name = "Month", 
    breaks = c(23,27,32,36, 40), 
    labels = c("Jun", "Jul", "Aug", "Sep", "Oct"))
  


```
##Phoebrain Practice##
 geom_point(aes(y = `Total Viscera Mass`), color = "red", alpha = 0.1) +
  geom_smooth(aes(y =`Shell Mass`), color = "blue") + 
  geom_smooth(aes(y = `Total Viscera Mass`), color = "red") +
  facet_wrap(~Year)+
  ylab("Mass")+
  scale_x_continuous(
    name = "Month", 
    breaks = c(23,27,32,36, 40), 
    labels = c("Jun", "Jul", "Aug", "Sep", "Oct"))

```

